Collective excitations in competing phases in two and three dimensions

### 2024-03-04 paper/main.tex
a
x

u

Paper stuff

data

stuff

tex

stuff

\newcommand{\vk}{\vec{k}}
\newcommand{\vl}{\vec{l}}
\newcommand{\vQ}{\vec{Q}}
\newcommand{\up}{\uparrow}
\newcommand{\down}{\downarrow}
\newcommand{\kplusQ}{\vk+\vQ}
\newcommand{\kminusQ}{-\vk-\vQ}

@@

We start with some operator set $\mathcal{B}$ that is complete with respect to commutation with a Hamiltonian $H$, 
i.e., any commutator $[H, A]$ for any $A \in \mathcal{B}$ can be represented by linear combinations of operators in $\mathcal{B}$.
We can then express any time-dependent operator from this set as
\begin{equation}
    a(t) = \sum_j c_j(t) A_j\,,
\end{equation}
where $c_j(t)$ captures the entire time dependence. 
Inserting this into the Heisenberg equation of motion and applying an operator scalar product $(A_i|\cdot)$ to both sides of the equation yields
\begin{equation}
    \label{eqn:heisenberg}
    \begin{aligned}
        \ddt a(t) = \sum_j \ddt c_j(t) A_j &= i \sum_j c_j(t) [H, A_j] \\
        \Rightarrow \sum_j \underbrace{(A_i | A_j)}_{\coloneqq \mN_{ij}} \ddt c_j(t) &= i \sum_j \underbrace{(A_i | [H, A_j])}_{\coloneqq \mM_{ij}} c_j(t) \\
        \Rightarrow \mN \ddt \vec{c}(t) &= i \mM \vec{c}(t)\,.
    \end{aligned}
\end{equation}
The matrices $\mN$ and $\mM$ contain all the dynamics and energetic properties of the system.
The former is referred to as a dynamical matrix while the latter is called a norm matrix.
The advantage is that one can now handle simple matrices rather than operators acting on an enormous Hilbert space.
The size of these matrices depends entirely on the set of operators.

@@

To study the collective excitations, we investigate various retarded Green's functions 
\begin{equation}
    G_{AB}^\text{ret} (t) = - i \langle [A(t), B] \rangle \Theta(t)\,,
\end{equation}
and their Fourier-transformed
\begin{equation}
    \label{eqn:standard_gf}
    G_{AB}(z = \omega + i0^+) = -i \int_0^{\infty} e^{izt} \langle [A(t), B] \rangle \mathrm{d}t\,,
\end{equation}
where $\Theta(t)$ is the Heaviside function.
Our previous choice of pseudo scalar product enables us to write down a matrix-valued Green's functions in terms of the aforementioned matrices
\begin{align}
    \label{eqn:green_function}
    \mathcal{G}(z &= \omega + i0^+) = \mN \frac{1}{-z \mN - \mM} \mN \\
        &\coloneqq  -\mN \mathcal{R}(z) \mN\,,
\end{align}
where we used the resolvent
\begin{equation}
    \label{eqn:resolvent}
    \mathcal{R}(z) \coloneqq  \frac{1}{z \mN + \mM}\,.
\end{equation}
The entries of this matrix $\mathcal{G}(z)$ are the Fourier-transformed Green's functions mentioned in \eqref{eqn:standard_gf} with respect to the operators from the initially chosen set.
\textcolor{red}{That means, if we index each operator in our set $A_i$, the matrix element $(i,j)$ of $\mathcal{G}$ is the Green's function
\begin{align}
    \mathcal{G}_{ij}(z = \omega +i0^+) &=  G_{A_i A_j^\dagger} (z) \nonumber \\
        &= -i \int_0^{\infty} \langle [A_i(t), A_j^\dagger(0)] \rangle e^{izt} \mathrm{d}t\,.
\end{align}}

In the following, we will derive this relation.
For that, consider the Fourier-transformed Green's function \eqref{eqn:standard_gf} with respect to two operators $A$ and $B^\dagger$ that are included in our operator set
\begin{align}
    G_{AB^\dagger} (z) &= - i \int_0^\infty \exp(i z t) \langle [ A(t), B^\dagger(0) ] \rangle \mathrm{d}t
\end{align}
Since $A$ and $B$ are included in the operator set, we can rewrite them as dot products of the vector
\begin{equation}
    \vec{A} \coloneqq  \left( A_1, A_2, \dots, A_N \right)\,, \quad
    \vec{A}^\dagger \coloneqq  \begin{pmatrix}
        A_1^\dagger \\ A_2^\dagger \\ \vdots \\ A_N^\dagger
    \end{pmatrix}\,,
\end{equation}
and some other complex-valued vector $\vec{a}$ which has its coefficients set in such a way,
that \textcolor{red}{$\vec{A} \cdot \vec{a}$} is equal to the desired operator. 
For instance, $\vec{a} \equiv \vec{e}_1$ in order to represent $A_1$.

Furthermore, assuming, that $\mN$ is invertible, we can easily solve the differential equation \eqref{eqn:heisenberg} to obtain
\begin{equation}
    \vec{c}(t) = \exp \left( i \mN^{-1} \mM t \right) \vec{c}(0)\,.
\end{equation}
This allows us to finally obtain
\begin{align}
    \label{eqn:green_derivation}
    G_{AB^\dagger} (z) &= i \int_0^\infty \exp(izt) \langle [ (\vec{A \cdot} \vec{b}(0))^\dagger, \vec{A} \cdot \vec{a}(t) ] \rangle\ \mathrm{d}t \nonumber \\
        &= i \int_0^\infty \exp(izt) \vec{b}^\dagger(0) \underbrace{\langle [ \vec{A}^\dagger, \vec{A} ] \rangle}_{\equiv \mN} \vec{a}(t) \mathrm{d}t \nonumber \\
        &= i  \int_0^\infty \vec{b}^\dagger(0) \mN \exp \left( i \left(\mN^{-1} \mM + z \right) t \right) \vec{a}(0) \mathrm{d}t  \nonumber \\
        &= - \vec{b}^\dagger(0) \left[ \mN \frac{1}{\mN^{-1} \mM + z} \right] \vec{a}(0) \nonumber \\
        &= - \vec{b}^\dagger(0) \left[ \mN \frac{1}{ \underbrace{\mM + z \mN}}_{\equiv \mathcal{R}(z)} \mN \right] \vec{a}(0)\,,
\end{align}
where $\vec{a}(0)$ and $\vec{b}(0)$ are chosen such, that their product with $\vec{A}$ yields $A(0)$ and $B(0)$, respectively.
The task at hand is finding the resolvent as $\mathcal{R}$.

To do that, we embrace this matrix structure in \eqref{eqn:xp_set} and assume that all matrix entries are real.
Then, with $\mathcal{R}|_X$ denoting the block of a matrix $\mathcal{R}$ corresponding to the $X$-operators, we can calculate
\begin{align}
    r_X (z) &\coloneqq \mathcal{R}(z)|_X = \left. \frac{1}{\mM + z \mN} \right\vert_X \nonumber \\
        &= \left[ \frac{1}{\mM} - \frac{1}{\mM} z \mN \frac{1}{\mM} + \frac{1}{\mM} z \mN \frac{1}{\mM} z \mN \frac{1}{\mM} - \cdots \right]_X \nonumber \\
        &= \left. \frac{1}{\mM} \sum_{j=0}^\infty \left( -z \mN \frac{1}{\mM} \right)^j \right\vert_X \,.
\end{align}
We may omit every second element of the sum, as we want to obtain the upper left block of the matrix, and $\mN$ always swaps the upper left and lower right block.
Thus, we obtain
\begin{align}
    r_X (z) &= \left. \frac{1}{\mM} \sum_{j=0}^\infty z^{2j} \left( \mN \frac{1}{\mM} \right)^{2j} \right\vert_X \nonumber \\
        &= \left. \frac{1}{\mM - z^2 \mN \mM^{-1} \mN} \right\vert_X \nonumber \\
        &= \frac{1}{\mathcal{K}_+ - z^2 \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger}\,.
\end{align}
Next, we define $\check{N} \coloneqq \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger$ for brevity.
Since $\mathcal{M}$ is positive semidefinite and block diagonal, its blocks $\mathcal{K}_\pm$ must be positive semidefinite as well.
Therefore, $\check{N}$ is also positive semidefinite and we can define its square root.
We obtain
\begin{equation}
    r_X (z) = -\check{N}^{-1/2} \frac{1}{z^2 - \check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2}} \check{N}^{-1/2}\,.
\end{equation}
The matrix $\check{M} \coloneqq \check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2}$ is positive semidefinite and Hermitian by the previous arguments.
If $\check{N}$ has a singular part, the inverse can be replaced by the pseudo-inverse.
The same calculation can be repeated by exchanging $X$ with $P$, yielding the relation for the remaining block.
Then, the definitions of the matrices above change in the following way
\begin{equation}
    \check{N}_P \coloneqq \mathcal{L}^\dagger \mathcal{K}_+^{-1} \mathcal{L}\,,\quad \check{M}_P \coloneqq \check{N}_P^{-1/2} \mathcal{K}_- \check{N}_P^{-1/2}\,.
\end{equation}
In conclusion, the Green's function is given by
\begin{subequations}
    \begin{align}
        G_{AB^\dagger} \vert_P (z) &= - \left[ \vec{b}^\dagger (0) \right]_P \mathcal{L}^\dagger r_X (z) \mathcal{L} \left[ \vec{a}(0) \right]_P\,, \\
        G_{AB^\dagger} \vert_X (z) &= - \left[ \vec{b}^\dagger (0) \right]_X \mathcal{L} r_P (z) \mathcal{L}^\dagger \left[ \vec{a}(0) \right]_X\,.
    \end{align}
\end{subequations}

@@

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/resolvent_overview_SC_CDW.pdf}
    \caption{Spectral functions with respect to the operators listed in \eqref{eqn:resolvent_bases}.
    The left column shows the results for the square lattice while the right column shows the results for the simple cubic lattice.
    The left edge of each plot is at $-0.05$ rather than $0$ in order to improve the visibility of the phase peak.
    The shaded area marks the two-particle continuum.
    For each plot, the parameters are set to $U=-2.5t$ and $N_\gamma = 6000$, while $V$ is varied according to the text in the individual panels.
    The panels show the spectral functions in the (a) SC phase and the (b and c) CDW phase, respectively.}
    \label{fig:resolvent_overview_SC}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/resolvent_overview_V0.pdf}
    \caption{Spectral functions with respect to the operators listed in \eqref{eqn:resolvent_bases}.
    The left column shows the results for the square lattice while the right column shows the results for the simple cubic lattice.
    The left edge of each plot is at $-0.05$ rather than $0$ in order to improve the visibility of the peaks at $\omega=0$.
    The shaded area marks the two-particle continuum.
    For each plot, the parameters are set to $U=-2.5t$, $V=0$, and $N_\gamma = 6000$.
    Since the ratio of $\Delta_\text{SC}$ to $\Delta_\text{CDW}$ may be chosen arbitrarily, we show in (a) the spectral functions for $\Delta_\text{CDW} = 0$ and in (b) for $\Delta_\text{SC} = 0$.}
    \label{fig:resolvent_overview_V0}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/resolvent_overview_AFM_CDW.pdf}
    \caption{Spectral functions with respect to the operators listed in \eqref{eqn:resolvent_bases}.
    The left column shows the results for the square lattice while the right column shows the results for the simple cubic lattice.
    The shaded area marks the two-particle continuum.
    For each plot, the parameters are set to $V=1t$ and $N_\gamma = 6000$, while $U$ is varied according to the text in the individual panels.
    The panels (a) and (c) show the spectral functions in the AFM phase, and the panels (b) and (d) in the CDW phase.}
    \label{fig:resolvent_overview_AFM}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/sc_peak_in_cdw.pdf}
    \caption{The upper panels (a) show double logarithmic plots of the positions $\omega_0$ of the peak in $\spectral{SC}$ in the CDW phase while the lower panels (b) show their respective weights $W_0 = e^{b}$.
    In the upper panels, the pink markers additionally show the peak position divided by the gap $\omega_0 / \Delta_\text{CDW}$.
    The left column shows the results for the square and the right column for the simple cubic lattice at $U=-2.5t$.
    The fits are linear, i.e., $y(V) = c \ln(V/t) + d$, and the parameters are printed on the individual panels.}
    \label{fig:sc_in_cdw_behavior}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/cdw_peak_in_sc.pdf}
    \caption{The panels (a) show plots of the positions $\omega_0$ of the peak in $\spectral{CDW}$ in the SC phase.
    Below that, in panels (b), their respective weights are depicted.
    The dashed lines in both rows show the position or weight computed for $V=0$ with $\Delta_\text{CDW} = 0$.
    The left column shows the results for the square and the right column for the simple cubic lattice at $U=-2.5t$.
    The left-most values are $V=-0.28t$ for the square lattice and $V=-0.34t$ for the simple cubic lattice.
    Choosing even smaller values results in $\mathcal{M}$ no longer being positive, i.e., the system is not in thermal equilibrium.
    \newline
    The lower panels (c) and (d) present the data again but with logarithmic scaling on the $V$-axis.
    Additionally, the peak positions and weights are given relative to their value at $V=0$.
    These data are fitted by $y(x) = a (\tanh (bx - c) + 1)$ with the parameters printed in the individual panel.
    The panels (a) and (b) feature the same fitting result.}
    \label{fig:cdw_in_sc_behavior}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/afm_cdw_peaks_overview.pdf}
    \caption{The upper panels (a) show plots of the positions $\omega_0$ of the peak in $\spectral{CDW}$ (orange lines) and $\spectral{AFM}$ (blue lines) close to the corresponding phase transition $U = zV$.
    The middle panels (b) depict the peak positions relative to the lower edge of the two-particle continuum $\omega_-$, while the bottom panels (c) show their respective weights.
    The shadings indicate the phase the system is in. Blue represents AFM while orange represents CDW.
    The left column shows the results for the square and the right column for the simple cubic lattice at $V=-1t$.
    Note the difference in the scales for the two lattices.}
    \label{fig:afm_cdw_peaks_overview}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/afm_cdw_peaks_details.pdf}
    \caption{The upper panels (a) show plots of the positions $\omega_0$ of the peak in $\spectral{AFM}$ relative to the lower edge of the two-particle continuum $\omega_-$.
    The lower panels (b) depict their respective weights $W_0$.
    The left column shows the results for the square and the right column for the simple cubic lattice at $V=-1t$.
    $U_0$ denotes the highest value of $U$ for which there is no peak below the two-particle continuum.
    For the square lattice, it is $U_0 = 3.5775t$ and for the simple cubic lattice, it is $U_0 = 4.89t$.
    The parameter range depicted here places the system within the CDW phase.
    All fits are linear, i.e., $y=ax + b$.}
    \label{fig:afm_cdw_peaks_details}
\end{figure*}

@@

Now, the task at hand is finding the inverse $1/(z^2 - \check{M})$ for all $z$.
This can be achieved using a Lanczos tridiagonalization and subsequently a continued fraction expansion in the form of

\begin{widetext}
\begin{equation}
    \left( z^2 - \begin{pmatrix}
        a_0 & b_1 & 0 & 0 & \cdots \\
        b_1 & a_1 & b_2 & 0 & \cdots \\
        0 & b_2 & a_2 & b_3 & \cdots \\
        0 & 0 & b_3 & a_3 & \cdots \\
        \vdots & \vdots & \vdots & \vdots & \ddots
    \end{pmatrix} \right)_{00}^{-1} = \dfrac{1}{z^2 - a_0 - \dfrac{b_1^2}{z^2 - a_1 - \dfrac{b_2^2}{ z^2 - a_2 - \hdots}}}\,\,
\end{equation}
\end{widetext}

where $a_i$ and $b_i$ are the Lanczos coefficients \cite{PettiforRecursion,ViswanathRecursion}.

Obtaining a diagonal block of $\mathcal{G}$ is easily achieved by setting $\vec{b}(0) = \vec{a}(0)$.
Obtaining an off-diagonal block requires linear combinations in the form of 
\begin{align}
    \vec{a}^\dagger \mathcal{G} \vec{b} &= \frac{1}{4} \Big\{ 
        \underbrace{\left[ (\vec{a} + \vec{b})^\dagger \mathcal{G} (\vec{a} + \vec{b}) - (\vec{a} - \vec{b})^\dagger \mathcal{G} (\vec{a} - \vec{b}) \right]}_{ = 2 \vec{a}^\dagger \mathcal{G} \vec{b} + 2 \vec{b}^\dagger \mathcal{G} \vec{a}} \\
    &-i \underbrace{\left[ (\vec{a} + i \vec{b})^\dagger \mathcal{G} (\vec{a} + i \vec{b}) - (\vec{a} - i \vec{b})^\dagger \mathcal{G} (\vec{a} - i \vec{b}) \right]}_{= 2i \vec{a}^\dagger \mathcal{G} \vec{b} - 2i \vec{b}^\dagger \mathcal{G} \vec{a}} \Big\}\,, \nonumber
\end{align}
i.e., one computes the four diagonal blocks and then simply adds them as described above.

In the single-band case, the coefficients approach the limit
\begin{equation}
    \label{eqn:inf_lanczos}
    a_\infty = \frac{\omega_+ + \omega_-}{2}\text{  and  } b_\infty = \frac{\omega_+ - \omega_-}{4}\,,
\end{equation}
where $\omega_\pm$ represents the upper and lower band edge, respectively \cite{PettiforRecursion}.
This behavior allows us to terminate the continued fraction, once the coefficients have converged close enough to these values, using the so-called square root terminator
\begin{equation}
    T(\omega) = \frac{1}{2b_\infty^2} \left( \omega - a_\infty \mp \sqrt{(\omega - a_\infty)^2 - 4 b_\infty^2} \right)\,,
\end{equation}
where the negative sign is to be chosen for $\omega - a_\infty > 2b_\infty$ and $|\omega - a_\infty| < 2b_\infty$ and the positive one otherwise.
It is to be noted, that the coefficients start to drop off from this limit as more and more Lanczos iterations are performed \cite{ViswanathRecursion}.
Therefore, we terminate the continued fraction with the aforementioned terminator at the coefficient pair that has the least deviation from $a_\infty$ and $b_\infty$.

@@

\section{Antisymmetry of the spectral functions}
\label{sec:antisymmetry_spectral}

>>>>

\newcommand{\vk}{\vec{k}}
\newcommand{\vl}{\vec{l}}
\newcommand{\vQ}{\vec{Q}}
\newcommand{\up}{\uparrow}
\newcommand{\down}{\downarrow}
\newcommand{\kplusQ}{\vk+\vQ}
\newcommand{\kminusQ}{-\vk-\vQ}
\newcommand{\im}{\mathrm{i}}

@@

We start with some operator set $\mathcal{B}$ that is complete with respect to commutation with a Hamiltonian $H$, 
i.e., any commutator $[H, A]$ for any $A \in \mathcal{B}$ can be represented by linear combinations of operators in $\mathcal{B}$.
We can then express any time-dependent operator from this set as
\begin{equation}
    \label{eqn:time_dependent_operator}
    a(t) = \sum_j c_j(t) A_j\,,
\end{equation}
where $c_j(t)$ captures the entire time dependence. 
Inserting this into the Heisenberg equation of motion and applying an operator scalar product $(A_i|\cdot)$ to both sides of the equation yields
\begin{equation}
    \label{eqn:heisenberg}
    \begin{aligned}
        \ddt a(t) = \sum_j \ddt c_j(t) A_j &= \im \sum_j c_j(t) [H, A_j] \\
        \Rightarrow \sum_j \underbrace{(A_i | A_j)}_{\coloneqq \mN_{ij}} \ddt c_j(t) &= \im \sum_j \underbrace{(A_i | [H, A_j])}_{\coloneqq \mM_{ij}} c_j(t) \\
        \Rightarrow \mN \ddt \vec{c}(t) &= \im \mM \vec{c}(t)\,.
    \end{aligned}
\end{equation}
The matrices $\mN$ and $\mM$ contain all the dynamics and energetic properties of the system.
The former is referred to as a dynamical matrix while the latter is called a norm matrix.
The advantage is that one can now handle simple matrices rather than operators acting on an enormous Hilbert space.
The size of these matrices depends entirely on the set of operators.

@@

To study the collective excitations, we investigate various retarded Green's functions 
\begin{equation}
    G_{AB}^\text{ret} (t) = - \im \langle [A(t), B] \rangle \Theta(t)\,,
\end{equation}
and their Fourier-transformed
\begin{equation}
    \label{eqn:standard_gf}
    G_{AB}(z = \omega + \im 0^+) = -\im \int_0^{\infty} e^{\im z t} \langle [A(t), B] \rangle \mathrm{d}t\,,
\end{equation}
where $\Theta(t)$ is the Heaviside function.
Our previous choice of pseudo scalar product enables us to write down a matrix-valued Green's functions in terms of the aforementioned matrices
\begin{align}
    \label{eqn:green_function}
    \mathcal{G}(z &= \omega + \im 0^+) = \mN \frac{1}{-z \mN - \mM} \mN \\
        &\coloneqq  -\mN \mathcal{R}(z) \mN\,,
\end{align}
where we used the resolvent
\begin{equation}
    \label{eqn:resolvent}
    \mathcal{R}(z) \coloneqq  \frac{1}{z \mN + \mM}\,.
\end{equation}
The entries of this matrix $\mathcal{G}(z)$ are the Fourier-transformed Green's functions mentioned in \eqref{eqn:standard_gf} with respect to the operators from the initially chosen set.
\textcolor{red}{That means, if we index each operator in our set $A_i$, the matrix element $(j,i)$ of $\mathcal{G}$ is the Green's function
\begin{align}
    \mathcal{G}_{ji}(z = \omega +\im 0^+) &=  G_{A_j A_i^\dagger} (z) \nonumber \\
        &= -\im \int_0^{\infty} \langle [A_j(t), A_i^\dagger(0)] \rangle e^{\im z t} \mathrm{d}t\,.
\end{align}}

In the following, we will derive this relation.
\textcolor{red}{For that, consider the Fourier-transformed Green's function above.
We rewrite both operators in the same manner as in \eqref{eqn:time_dependent_operator} to obtain
\begin{align}
    G_{AB^\dagger} (z) &= \im \int_0^\infty e^{\im z t} \langle [ A_i^\dagger(0), A_j(t) ] \rangle \mathrm{d}t \nonumber \\
        &= \im \int_0^\infty e^{\im z t} \langle [ \sum_m a_{j,m}^*(0) A_m^\dagger, \sum_n a_{i,n}(t) A_n ] \rangle \mathrm{d}t \nonumber \\
        &= \im \int_0^\infty \sum_{mn} e^{\im z t} a_{j,m}^*(0) \underbrace{\langle [ A_m^\dagger, A_n ] \rangle}_{\equiv \mN_{mn}} a_{i,n}(t) \mathrm{d}t \nonumber \\
        &= \im \int_0^\infty e^{\im z t} \vec{a}_{j}^\dagger (0) \mN \vec{a}_{i}(t) \mathrm{d}t\,.
\end{align}
Furthermore, assuming, that $\mN$ is invertible, we can easily solve the differential equation \eqref{eqn:heisenberg} to obtain
\begin{equation}
    \vec{a}_{i}(t) = \exp \left( \im \mN^{-1} \mM t \right) \vec{a}_{i}(0)\,.
\end{equation}
This allows us to finally obtain
\begin{align}
    \label{eqn:green_derivation}
    G_{AB^\dagger} (z) &= \im  \int_0^\infty \vec{a}_{j}^\dagger (0) \mN \exp \left( \im \left(\mN^{-1} \mM + z \right) t \right) \vec{a}_{i}(0) \mathrm{d}t  \nonumber \\
        &= - \vec{a}_{j}^\dagger (0) \left[ \mN \frac{1}{\mN^{-1} \mM + z} \right] \vec{a}_{i}(0) \nonumber \\
        &= - \vec{a}_{j}^\dagger (0) \left[ \mN \frac{1}{ \underbrace{\mM + z \mN}}_{\equiv \mathcal{R}(z)} \mN \right] \vec{a}_{i}(0)\,,
\end{align}
where $\vec{a}_i(0)$ and $\vec{a}_j^\dagger (0)$ need to be chosen such, that $A_i = \sum_n a_{i,n} A_n$ and $A_j^\dagger = \sum_n a_{j,n}^* A_n^\dagger$ hold.
The task at hand is finding the resolvent as $\mathcal{R}$.}

To do that, we embrace this matrix structure in \eqref{eqn:xp_set} and assume that all matrix entries are real \textcolor{red}{so that $\kappa_ij = \Lambda_{\pm,ij} = 0$ holds}.
Then, with $\mathcal{R}|_X$ denoting the block of a matrix $\mathcal{R}$ corresponding to the $X$-operators, we can calculate
\begin{align}
    r_X (z) &\coloneqq \mathcal{R}(z)|_X = \left. \frac{1}{\mM + z \mN} \right\vert_X \nonumber \\
        &= \left[ \frac{1}{\mM} - \frac{1}{\mM} z \mN \frac{1}{\mM} + \frac{1}{\mM} z \mN \frac{1}{\mM} z \mN \frac{1}{\mM} - \cdots \right]_X \nonumber \\
        &= \left. \frac{1}{\mM} \sum_{j=0}^\infty \left( -z \mN \frac{1}{\mM} \right)^j \right\vert_X \,.
\end{align}
We may omit every second element of the sum, as we want to obtain the upper left block of the matrix, and $\mN$ always swaps the upper left and lower right block.
Thus, we obtain
\begin{align}
    r_X (z) &= \left. \frac{1}{\mM} \sum_{j=0}^\infty z^{2j} \left( \mN \frac{1}{\mM} \right)^{2j} \right\vert_X \nonumber \\
        &= \left. \frac{1}{\mM - z^2 \mN \mM^{-1} \mN} \right\vert_X \nonumber \\
        &= \frac{1}{\mathcal{K}_+ - z^2 \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger}\,.
\end{align}
Next, we define $\check{N} \coloneqq \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger$ for brevity.
Since $\mathcal{M}$ is positive semidefinite and block diagonal, its blocks $\mathcal{K}_\pm$ must be positive semidefinite as well.
Therefore, $\check{N}$ is also positive semidefinite and we can define its square root.
We obtain
\begin{equation}
    \label{eqn:rx}
    r_X (z) = -\check{N}^{-1/2} \frac{1}{z^2 - \check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2}} \check{N}^{-1/2}\,.
\end{equation}
The matrix $\check{M} \coloneqq \check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2}$ is positive semidefinite and Hermitian by the previous arguments.
If $\check{N}$ has a singular part, the inverse can be replaced by the Moore-Penrose pseudo-inverse.
\textcolor{red}{A possible way to compute this is to diagonalize $\check{N} = \mathcal{V} \mathcal{D} \mathcal{V}^\dagger$, with a diagonal matrix $\mathcal{D}$ and a unitary transformation matrix $\mathcal{V}$.
Now, the pseudo-inverse can be defined as
\begin{equation}
    \check{N}^+ \coloneqq \mathcal{V} \mathcal{D}^+ \mathcal{V}^\dagger
\end{equation}
with
\begin{equation}
    (\mathcal{D}^+)_{ij} = \begin{cases}
        1/\mathcal{D}_{ij} & \mathcal{D}_{ij} \neq 0 \\ 0 & \text{otherwise}
    \end{cases}\,.
\end{equation}
Note, that this definition yields the standard inverse for invertible matrices.}

The same calculation can be repeated by exchanging $X$ with $P$, yielding the relation for the remaining block.
Then, the definitions of the matrices above change in the following way
\begin{equation}
    \check{N}_P \coloneqq \mathcal{L}^\dagger \mathcal{K}_+^{-1} \mathcal{L}\,,\quad \check{M}_P \coloneqq \check{N}_P^{-1/2} \mathcal{K}_- \check{N}_P^{-1/2}\,.
\end{equation}
In conclusion, the Green's function is given by
\begin{subequations}
    \begin{align}
        G_{AB^\dagger} \vert_P (z) &= - \left[ \vec{b}^\dagger (0) \right]_P \mathcal{L}^\dagger r_X (z) \mathcal{L} \left[ \vec{a}(0) \right]_P\,, \\
        G_{AB^\dagger} \vert_X (z) &= - \left[ \vec{b}^\dagger (0) \right]_X \mathcal{L} r_P (z) \mathcal{L}^\dagger \left[ \vec{a}(0) \right]_X\,,
    \end{align}
\end{subequations}
\textcolor{red}{where $r_X (z)$ is given in \eqref{eqn:rx} and 
\begin{equation}
    r_P (z) = -\check{N}_P^{-1/2} \frac{1}{z^2 - \check{M}_P} \check{N}_P^{-1/2}\,.
\end{equation}}

Now, the task at hand is finding the inverse $1/(z^2 - \check{M})$ for all $z$.
This can be achieved using a Lanczos tridiagonalization and subsequently, a continued fraction expansion in terms of the Lanczos coefficients $a_i$ and $b_i$ \cite{PettiforRecursion,ViswanathRecursion}.
In the single-band case, the coefficients approach the limit
\begin{equation}
    \label{eqn:inf_lanczos}
    a_\infty = \frac{\omega_+ + \omega_-}{2}\text{  and  } b_\infty = \frac{\omega_+ - \omega_-}{4}\,,
\end{equation}
where $\omega_\pm$ represents the upper and lower band edge, respectively \cite{PettiforRecursion}.
This behavior allows us to terminate the continued fraction, once the coefficients have converged close enough to these values, using the so-called square root terminator
\begin{equation}
    T(\omega) = \frac{1}{2b_\infty^2} \left( \omega - a_\infty \mp \sqrt{(\omega - a_\infty)^2 - 4 b_\infty^2} \right)\,,
\end{equation}
where the negative sign is to be chosen for $\omega - a_\infty > 2b_\infty$ and $|\omega - a_\infty| < 2b_\infty$ and the positive one otherwise.
It is to be noted, that the coefficients start to drop off from this limit as more and more Lanczos iterations are performed \cite{ViswanathRecursion}.
Therefore, we terminate the continued fraction with the aforementioned terminator at the coefficient pair that has the least deviation from $a_\infty$ and $b_\infty$.

@@

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/resolvent_overview_SC_CDW.pdf}
    \caption{Spectral functions with respect to the operators listed in Eq.~\eqref{eqn:resolvent_bases}.
    The left column shows the results for the square lattice while the right column shows the results for the simple cubic lattice.
    The left edge of each plot is at $-0.05$ rather than $0$ in order to improve the visibility of the phase peak.
    The shaded area marks the two-particle continuum.
    For each plot, the parameters are set to $U=-2.5t$ and $N_\gamma = 6000$, while $V$ is varied, see the legend.
    The panels show the spectral functions in the (a) SC phase and the (b and c) CDW phase, respectively.}
    \label{fig:resolvent_overview_SC}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/resolvent_overview_V0.pdf}
    \caption{\textcolor{red}{Same as \autoref{fig:resolvent_overview_SC}, except that we set $V=0$, i.e, SC and CDW order coexist.}
    Since the ratio of $\Delta_\text{SC}$ to $\Delta_\text{CDW}$ may be chosen arbitrarily, we show in (a) the spectral functions for $\Delta_\text{CDW} = 0$ and in (b) for $\Delta_\text{SC} = 0$.}
    \label{fig:resolvent_overview_V0}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/resolvent_overview_AFM_CDW.pdf}
    \caption{\textcolor{red}{Same as \autoref{fig:resolvent_overview_SC}, except that we are now looking at the system near the AFM-CDW phase transition.
    For the square lattice, we set $V=1.2t$ and for the simple cubic lattice, we set $V=0.8t$. This choice yields the phase boundary at $U=4.8t$ for both lattices.}
    The panels (a) and (c) show the spectral functions in the AFM phase, and the panels (b) and (d) in the CDW phase.}
    \label{fig:resolvent_overview_AFM}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/sc_peak_in_cdw.pdf}
    \caption{The upper panels (a) show double logarithmic plots of the positions $\omega_0$ of the peak in $\spectral{SC}$ in the CDW phase while the lower panels (b) show their respective weights $W_0$.
    In the upper panels, the pink markers additionally show the peak position divided by the gap $\omega_0 / \Delta_\text{CDW}$.
    The left column shows the results for the square and the right column for the simple cubic lattice at $U=-2.5t$.
    \textcolor{red}{The fits are linear, i.e., $y(V) = e^d V^c$, and the parameters are printed on the individual panels.}}
    \label{fig:sc_in_cdw_behavior}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/cdw_peak_in_sc.pdf}
    \caption{\textcolor{red}{Same as \autoref{fig:sc_in_cdw_behavior} but for the peak in $\spectral{CDW}$ within the SC phase, i.e., $V<0$.}}
    \label{fig:cdw_in_sc_behavior}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/afm_cdw_peaks_overview.pdf}
    \caption{\textcolor{red}{The upper panels (a) show plots of the positions $\omega_0$ of the peak in $\spectral{CDW}$ (orange lines) and $\spectral{l.AFM}$ (blue lines) close to the corresponding phase transition $U = 4.8t$.
    Again, we chose $V=1.2t$ for the square lattice (left column) and $V=0.8t$ for the simple cubic lattice (right column) respectively.
    The middle panels (b) depict the peak positions relative to the lower edge of the two-particle continuum $\omega_-$, while the bottom panels (c) show their respective weights.}
    The shadings indicate the phase the system is in. Blue represents AFM while orange represents CDW.
    Note the difference in the scales for the two lattices.}
    \label{fig:afm_cdw_peaks_overview}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=.98\textwidth]{plots/afm_cdw_peaks_details.pdf}
    \caption{The upper panels (a) show double logarithmic plots of the positions $\omega_0$ of the peak in $\spectral{l.AFM}$ relative to the lower edge of the two-particle continuum $\omega_-$.
    The lower panels (b) depict their respective weights $W_0$.
    Again, we chose $V=1.2t$ for the square lattice (left column) and $V=0.8t$ for the simple cubic lattice (right column) respectively.
    $U_0$ denotes the highest value of $U$ for which there is no peak below the two-particle continuum.
    For the square lattice, it is $U_0 \approx 3.372t$ and for the simple cubic lattice, it is $U_0 \approx 4.479t$.
    The parameter range depicted here places the system within the CDW phase.
    The fits are of the same fashion as in \autoref{fig:sc_in_cdw_behavior}.}
    \label{fig:afm_cdw_peaks_details}
\end{figure*}

@@

@@
